---
title: "Continuous and mixed-type comparisons in reclin2"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Continuous and mixed-type comparisons}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

The classic Fellegi-Sunter model in `reclin2` (see the *Introduction* vignette)
compares records on *binary* agreement: for each linkage key a pair either
agrees or it does not. This works very well when the matching records agree
*exactly* on their keys -- names after standardisation, postcodes, dates of
birth that were copied without error.

It throws information away, however, when the keys are *numeric* and the
matching records are *close but rarely exactly equal*: ages, incomes,
coordinates, dates measured to the day, sensor readings. Two records for the
same person might have ages 41 and 42, or incomes 30 000 and 30 150 -- clearly
similar, but a binary comparator scores both disagreements identically to a
comparison of 41 against 88.

This vignette covers two estimators that use the full numeric information:

- `problink_em_gamma()` -- a **continuous** Fellegi-Sunter model. It models the
  comparison *distance* (e.g. `|age_x - age_y|`) with a two-component
  hurdle-gamma mixture: a small mean distance for matches, a large one for
  non-matches.
- `problink_em_mixed()` -- a **mixed-type** model in which each field can be
  binary, categorical (ordinal agreement levels) or continuous, combined under
  the usual conditional-independence assumption.

Both produce exactly the same output columns as the standard `problink_em`, so
`predict()`, `select_threshold()`, `select_n_to_m()`, `link()` and friends work
unchanged.

The estimators follow the extension of the Fellegi-Sunter model to numeric and
mixed-type comparisons proposed by Vo, Chauvet, Happe, Oger, Paquelet and
Garès (2023), "Extending the Fellegi-Sunter record linkage model for mixed-type
data", *Computational Statistics & Data Analysis* **179**:107656.

> **When does this help?** When matches agree *exactly* on their keys, plain
> binary Fellegi-Sunter already does well and is simpler -- reach for it first.
> The continuous and mixed models earn their keep when matches are
> *close-but-rarely-exact*, which is precisely where binary comparison loses
> information.

```{r libs, message = FALSE}
library(reclin2)
library(data.table)
```

## Theory: the Fellegi-Sunter model and its extensions

### The classic Fellegi-Sunter recap

For a candidate pair $(i, j)$ formed from records in two files $A$ and $B$, the
Fellegi-Sunter model introduces a **latent match indicator**
$g_{ij} \in \{0, 1\}$: $g_{ij} = 1$ if the pair is a true *match* (the two
records describe the same entity, the set $M$) and $g_{ij} = 0$ if it is a
*non-match* (the set $U$). What we observe instead is a **comparison vector**
$\gamma_{ij} = (\gamma_{ij}^{(1)}, \dots, \gamma_{ij}^{(K)})$ summarising how the
pair agrees on each of $K$ linkage fields.

The model is a two-component mixture governed by:

- the **match prevalence** $p_M = \Pr(g_{ij} = 1)$, the overall probability that
  a random pair is a match;
- the **m-parameters** $\Pr(\gamma \mid M)$, the distribution of the comparison
  vector among true matches;
- the **u-parameters** $\Pr(\gamma \mid U)$, the distribution among non-matches.

Under the standard **conditional independence** assumption, the comparison
fields are independent given match status, so the joint comparison densities
factorise across fields,

$$
\Pr(\gamma_{ij} \mid M) = \prod_{k=1}^{K} \Pr\!\big(\gamma_{ij}^{(k)} \mid M\big),
\qquad
\Pr(\gamma_{ij} \mid U) = \prod_{k=1}^{K} \Pr\!\big(\gamma_{ij}^{(k)} \mid U\big).
$$

Linkage decisions are driven by the **posterior match probability** obtained
from Bayes' rule,

$$
q_{ij}
= \Pr(M \mid \gamma_{ij})
= \frac{p_M\,\Pr(\gamma_{ij}\mid M)}
       {p_M\,\Pr(\gamma_{ij}\mid M) + (1-p_M)\,\Pr(\gamma_{ij}\mid U)},
$$

or, equivalently, by the **matching weight** -- the log-likelihood ratio that
accumulates additively over fields,

$$
w_{ij}
= \log\frac{\Pr(\gamma_{ij}\mid M)}{\Pr(\gamma_{ij}\mid U)}
= \sum_{k=1}^{K}
  \log\frac{\Pr(\gamma_{ij}^{(k)}\mid M)}{\Pr(\gamma_{ij}^{(k)}\mid U)}.
$$

In the *classic* model each $\gamma^{(k)}$ is **binary** (agree/disagree), so
each field contributes the familiar $m_k = \Pr(\text{agree}\mid M)$ and
$u_k = \Pr(\text{agree}\mid U)$. The two estimators below keep this exact mixture
skeleton -- same $p_M$, same posterior $q_{ij}$, same additive weight -- but
**replace the per-field densities** so that numeric and mixed comparison vectors
can be modelled directly (Vo et al., 2023).

### Continuous fields: the hurdle-gamma model

For a continuous key the comparison is no longer agree/disagree but a
**distance** $\gamma = |x - y| \ge 0$, where $\gamma = 0$ is an exact match and
larger values mean greater dissimilarity (this is exactly what `cmp_distance()`
returns). Distances are non-negative and have an *atom at zero* -- exact
agreements do happen for rounded data -- so a plain gamma density is not enough.
`problink_em_gamma()` models each field/component with a **hurdle-gamma**
density,

$$
f(\gamma;\, p_0, \alpha, \beta) =
\begin{cases}
p_0, & \gamma = 0,\\[4pt]
(1 - p_0)\,\operatorname{dgamma}(\gamma;\ \text{shape}=\alpha,\ \text{scale}=\beta),
  & \gamma > 0,
\end{cases}
$$

with a point mass $p_0$ on an exact match ("hurdle") and a gamma density for the
positive distances. Each of the two components (match $C=M$, non-match $C=U$) has
its own $(p_{0,C}, \alpha_C, \beta_C)$; the gamma **mean distance** is
$\alpha_C \beta_C$, small for matches and large for non-matches. The pair-level
density used in the mixture is the product over fields,
$\Pr(\gamma \mid C) = \prod_k f(\gamma^{(k)};\, p_{0,C}^{(k)}, \alpha_C^{(k)}, \beta_C^{(k)})$.

The parameters are fitted by EM. The **E-step** computes the posterior
$q_{ij} = \Pr(M \mid \gamma_{ij})$ with the Bayes formula above, using these
hurdle-gamma densities. The **M-step** maximises the expected complete-data
log-likelihood, giving, with weights $q_{ij}$ (match) and $1 - q_{ij}$
(non-match):

- match prevalence (overall mean posterior),
  $$
  p_M = \frac{1}{N}\sum_{ij} q_{ij};
  $$
- hurdle point mass (weighted fraction of exact zeros), for each component and
  field,
  $$
  p_0 = \frac{\sum_{ij} q_{ij}\,\mathbf{1}[\gamma_{ij}=0]}{\sum_{ij} q_{ij}};
  $$
- gamma **shape** $\alpha$, the root of the weighted score equation
  $$
  -\log\beta + \frac{\sum w\,\log\gamma}{\sum w} - \psi(\alpha) = 0,
  $$
  where $\psi$ is the digamma function and the weights are
  $w = q\,\mathbf{1}[\gamma > 0]$ for the match component (and
  $w = (1-q)\,\mathbf{1}[\gamma > 0]$ for the non-match component) -- i.e. only
  the positive distances enter the gamma update;
- gamma **scale** $\beta$ (closed form given $\alpha$),
  $$
  \beta = \frac{\sum w\,\gamma}{\alpha \sum w}.
  $$

When the data contain **no exact agreements** the numerator of every $p_0$
update is zero, so $p_0 \to 0$ (numerically, its floor) and the hurdle switches
off: the model collapses to an ordinary two-component gamma mixture over the
distances. This is the correct maximum-likelihood behaviour, not a failure --
matches are still recovered from the separation of the two gamma components.
`problink_em_gamma()` detects this regime and reports the hurdle as *inactive*.

### Mixed-type fields: a product of per-field blocks

Real linkage keys mix kinds: a sex indicator, a region code, an age, an income.
Under the same conditional-independence factorisation, each field contributes its
**own block** to $\Pr(\gamma \mid M)$ and $\Pr(\gamma \mid U)$ according to its
type, and the EM skeleton is unchanged -- only the per-field M-step differs (Vo
et al., 2023):

- **binary (Bernoulli).** $\gamma \in \{0, 1\}$; the block is
  $m^{\gamma}(1-m)^{1-\gamma}$ with $m = \Pr(\text{agree}\mid C)$. Closed-form
  M-step:
  $$
  m = \frac{\sum_{ij} q_{ij}\,\gamma_{ij}}{\sum_{ij} q_{ij}}
  $$
  for the match component (and the analogous $(1-q)$-weighted update for the
  non-match component $u$).
- **categorical (multinomial, $L$ levels).** $\gamma \in \{0, \dots, L-1\}$ is an
  ordinal agreement level (e.g. from `cmp_levels()`); the block is
  $\prod_{s} m_s^{\mathbf{1}[\gamma = s]}$ with $\sum_s m_s = 1$. Closed-form
  M-step (weighted relative frequency of each level):
  $$
  m_s = \frac{\sum_{ij} q_{ij}\,\mathbf{1}[\gamma_{ij}=s]}{\sum_{ij} q_{ij}},
  \qquad s = 0, \dots, L-1.
  $$
- **continuous.** the hurdle-gamma block from the previous section, with its
  $(p_0, \alpha, \beta)$ updated exactly as in `problink_em_gamma()`.

So the binary and categorical updates are closed form, while the continuous
update is the gamma score-equation solve. A model in which *every* field is
continuous reduces **exactly** to `problink_em_gamma()`.

## Comparators that return distances

Two new comparators feed these models:

- `cmp_distance(scale, transform)` returns the scaled absolute distance
  `|x - y| / scale`; **0 means an exact match** and larger values mean greater
  dissimilarity. `cmp_absdist()` is the shorthand `cmp_distance(scale = 1)`.
- `cmp_levels(breaks)` discretises `|x - y|` into ordinal *agreement levels*
  using cut points. `cmp_levels(c(0, 3))` gives level `0` for an exact match,
  `1` for a distance in `(0, 3]` and `2` for a distance above `3`.

```{r comparators}
cmp_distance()(c(41, 30000, 5), c(42, 30150, 5))   # distances; 0 = exact
cmp_levels(c(0, 3))(c(0, 2, 10), c(0, 0, 0))       # ordinal levels 0/1/2
```

Both comparators are *polymorphic*: called with a single argument they fall
back to a binary "exact match" result, so they can also be handed to the
classic `problink_em`.

## Continuous linkage with `problink_em_gamma`

We simulate a register `A` and a second file `B` that contains 80 of the same
people, each measured on three numeric keys with a small amount of error. The
errors are tiny relative to the spread of the keys, so matching records are
*close*, but -- being continuous -- essentially **never exactly equal**.

```{r cont-sim}
set.seed(2025)
nA <- 400L; nB <- 80L
A <- data.table(
  id = 1:nA,
  x1 = runif(nA, 0, 100),
  x2 = runif(nA, 0, 100),
  x3 = runif(nA, 0, 100)
)
B <- A[1:nB]
B[, x1 := x1 + rnorm(nB, 0, 1)]
B[, x2 := x2 + rnorm(nB, 0, 1)]
B[, x3 := x3 + rnorm(nB, 0, 1)]

# Not a single exact agreement, even though all 80 are true matches:
sum(A$x1[1:nB] == B$x1)
```

We generate all pairs and compare them with `cmp_distance()`. Because the keys
contain errors we do not block on them; in a real application you would block on
a separate, reliable key.

```{r cont-pairs}
pairs <- pair(A, B)
pairs <- compare_pairs(pairs, on = c("x1", "x2", "x3"),
  default_comparator = cmp_distance())
print(pairs)
```

The comparison columns now hold distances rather than `TRUE`/`FALSE`. We fit the
continuous model with a formula, exactly as with `problink_em`:

```{r cont-fit}
model <- problink_em_gamma(~ x1 + x2 + x3, data = pairs)
print(model)
```

The printout gives, per field and per component (match vs non-match): the hurdle
mass `p0` (the probability of an *exact* zero distance) and the gamma `shape`
and `scale`. Read off the **mean distance** as `shape * scale`: for a match it
is small (the keys nearly agree), for a non-match it is large.

Two diagnostics are worth noting:

- The **match prevalence** `p` is re-estimated by the EM; its starting value
  `p0` is derived from the data (see *Notes* below).
- The **hurdle point mass** is reported as *inactive* here: there are no exact
  matches in this purely continuous data, so each `p0` collapses to its floor
  `1/N` and the model becomes a plain two-component gamma mixture. That is the
  correct maximum-likelihood behaviour, not a failure -- the matches are still
  recovered from the separation of the two gamma components.

We score and select exactly as in the standard workflow. `predict(type =
"mpost")` adds the posterior match probability; `select_threshold` keeps the
pairs above a cut-off. We then attach the truth (we know it here because `id`
identifies the same person in both files) and evaluate.

```{r cont-eval}
pairs <- predict(model, pairs, type = "mpost", add = TRUE)
pairs <- select_threshold(pairs, "selected", "mpost", threshold = 0.5)
pairs <- compare_vars(pairs, "truth", on_x = "id", on_y = "id")

table(truth = pairs$truth, selected = pairs$selected)
```

```{r cont-tprppv}
TP <- sum(pairs$truth & pairs$selected)
FP <- sum(!pairs$truth & pairs$selected)
FN <- sum(pairs$truth & !pairs$selected)
c(TPR = TP / (TP + FN), PPV = TP / (TP + FP))
```

The continuous model recovers almost all of the 80 matches with few false
links. As in the introduction, forcing one-to-one linkage with `select_n_to_m`
(which also accepts the `weights` column) does even better:

```{r cont-ntom}
pairs <- predict(model, pairs, type = "weights", add = TRUE)
pairs <- select_n_to_m(pairs, "ntom", "weights", threshold = 0)
table(truth = pairs$truth, ntom = pairs$ntom)
```

### Why not just use binary Fellegi-Sunter here?

On this data the keys never agree exactly, so a binary comparator marks *every*
pair as a disagreement on *every* field and the classic EM has nothing to
separate matches from non-matches:

```{r cont-binary}
pb <- pair(A, B)
pb <- compare_pairs(pb, on = c("x1", "x2", "x3"),
  default_comparator = cmp_identical())
# binary FS cannot fit when nothing agrees exactly; the warnings are expected
mb <- suppressWarnings(problink_em(~ x1 + x2 + x3, data = pb))
pb <- predict(mb, pairs = pb, type = "mpost", add = TRUE)
pb <- select_threshold(pb, "selected", "mpost", threshold = 0.5)
pb <- compare_vars(pb, "truth", on_x = "id", on_y = "id")
c(selected = sum(pb$selected), found = sum(pb$truth & pb$selected))
```

It selects nothing: the binary model is the wrong tool for purely continuous
keys.

### The pure-continuous regime and `nstart`

When there are *no* exact matches at all (as above), the two-component gamma
mixture can occasionally settle on a poor local optimum from a single start.
Running several starts and keeping the best log-likelihood makes the fit much
more robust. Use `nstart >= 10` in this regime; with `nstart = 1` the estimator
is fully deterministic, and `nstart > 1` is reproducible under `set.seed()`:

```{r cont-nstart}
set.seed(1)
model_ms <- problink_em_gamma(~ x1 + x2 + x3, data = pairs, nstart = 10)
model_ms$loglik >= model$loglik       # multistart is never worse
```

## Mixed-type linkage with `problink_em_mixed`

Real linkage keys are rarely all of one kind. `problink_em_mixed` lets each
field be **binary**, **categorical** (ordinal agreement levels) or
**continuous**, and fits them jointly. We build a file with one of each:

- `sex` -- a binary field, compared with `cmp_identical()`;
- `reg` -- a region code with many values; we compare it with `cmp_levels(c(0,
  3))` to get ordinal agreement (exact / close / far);
- `age` and `inc` -- continuous, compared with `cmp_distance()`.

```{r mixed-sim}
set.seed(101)
nA <- 400L; nB <- 80L
A <- data.table(
  id  = 1:nA,
  sex = sample(0:1, nA, replace = TRUE),
  reg = sample(1:50, nA, replace = TRUE),
  age = round(runif(nA, 18, 85)),
  inc = round(rgamma(nA, shape = 2, scale = 20000))
)
B <- A[1:nB]
ix <- sample(nB, round(nB * 0.05)); B$sex[ix] <- 1 - B$sex[ix]
ix <- sample(nB, round(nB * 0.10)); B$reg[ix] <- sample(1:50, length(ix), replace = TRUE)
B[, age := age + rbinom(nB, 1, 0.3) * sample(1:4, nB, replace = TRUE)]
B[, inc := inc + round(rnorm(nB, 0, 800))]

pairs <- pair(A, B)
pairs <- compare_pairs(pairs, on = c("sex", "reg", "age", "inc"),
  comparators = list(
    sex = cmp_identical(),     # binary
    reg = cmp_levels(c(0, 3)), # ordinal: 0 exact, 1 within 3, 2 far
    age = cmp_distance(),      # continuous distance
    inc = cmp_distance()       # continuous distance
  ))
print(pairs)
```

We fit with `problink_em_mixed`. It **auto-detects** the comparison type of each
field and reports it via a message:

- a column whose values are all in `{0, 1}` -> `binary`;
- non-negative integers with few distinct values -> `categorical`;
- anything else -> `continuous`.

```{r mixed-fit}
model <- problink_em_mixed(~ sex + reg + age + inc, data = pairs)
print(model)
```

The printout shows one block per field: `m`/`u` agreement probabilities for the
binary field, the per-level `m`/`u` probabilities for the categorical field, and
the hurdle-gamma `p0`/`shape`/`scale` for each continuous field. The continuous
`age` block has its hurdle mass `p0` *active*: ages do sometimes agree exactly,
and the model uses that.

Scoring, selection and evaluation are identical to every other model in the
package:

```{r mixed-eval}
pairs <- predict(model, pairs, type = "mpost", add = TRUE)
pairs <- select_threshold(pairs, "selected", "mpost", threshold = 0.5)
pairs <- compare_vars(pairs, "truth", on_x = "id", on_y = "id")

table(truth = pairs$truth, selected = pairs$selected)
```

```{r mixed-tprppv}
TP <- sum(pairs$truth & pairs$selected)
FP <- sum(!pairs$truth & pairs$selected)
FN <- sum(pairs$truth & !pairs$selected)
c(TPR = TP / (TP + FN), PPV = TP / (TP + FP))
```

### Overriding the inferred types

Auto-detection has one genuinely ambiguous case: an *integer-valued distance*
with few distinct values (for example a date lag binned to `0, 1, 2, ...`)
detects as `categorical`, even though you might want it treated as a continuous
distance. Pass `types =` to override any field; unnamed fields keep their
detected type:

```{r mixed-override}
model2 <- problink_em_mixed(~ sex + reg + age + inc, data = pairs,
  types = list(reg = "continuous"))
model2$types
```

## Under the hood: estimation and numerics

This section maps the theory above onto what the implementation actually does.
The continuous estimator lives in `R/problink_em_gamma.R`
(`problink_em_gamma()`), the mixed-type estimator in `R/problink_em_mixed.R`
(`problink_em_mixed()`), the comparators in `R/comparators_distance.R`
(`cmp_distance()`, `cmp_levels()`), and scoring in the two `predict.*` methods.

**E-step on the log scale (log-sum-exp).** With many fields the raw density
products $\prod_k f(\gamma^{(k)})$ underflow to zero in double precision, which
would turn the posterior into $0/0$. The mixed estimator therefore accumulates
**per-field log-densities** and combines the two weighted components with a
log-sum-exp,
$$
q = \frac{\exp(a)}{\exp(a) + \exp(b)},\quad
a = \log p_M + \sum_k \log f_M(\gamma^{(k)}),\quad
b = \log(1-p_M) + \sum_k \log f_U(\gamma^{(k)}),
$$
evaluated as $q = \exp\!\big(a - m - \log(e^{a-m} + e^{b-m})\big)$ with
$m = \max(a, b)$. This is numerically what the Bayes posterior $q_{ij}$ in the
theory computes, just done safely.

**Data-driven initialisation.** Good starts matter for mixture EM, so the
defaults are derived from the data rather than fixed:

- the match-prevalence start is blocking-aware,
  `p0 = min(nA, nB) / nrow(pairs)` -- the largest plausible match rate (every
  record in the smaller file matches). Because `nrow(pairs)` already reflects any
  blocking you applied, the start adapts automatically; it is re-estimated by the
  EM. When the source-table sizes are unavailable (e.g. the
  `comparison_matrix =` interface) it falls back to `0.05`, clamped to
  `[1/N, 0.5]`.
- the hurdle point mass `p0M` for each continuous field is seeded from the
  *observed* exact-zero fraction `z_k = mean(distance == 0)`, inflated by the
  match prevalence (the zeros that occur are mostly contributed by true matches).
  With no exact matches `z_k = 0`, so `p0M -> 1/N` and the hurdle starts (and
  stays) inactive.
- the gamma `shape`/`scale` are seeded by splitting each field's *positive*
  distances into a small "match" tail (the smallest ones) and a "non-match" bulk
  (the rest) and fitting a gamma to each by method of moments (or
  `EnvStats::egamma` when `use_envstats = TRUE`).

**Robustness and numerical hardening.** The updates are guarded so the EM can
never error or emit `NaN`:

- every probability is clamped to $[1/N,\, 1 - 1/N]$ and every M-step
  denominator is floored away from zero, so a near-collapsed component cannot
  produce $0/0$;
- the gamma **shape** solver degrades gracefully through a chain:
  `uniroot` -> `nlminb` -> weighted method-of-moments -> previous value, each
  stage validated to return a finite positive shape (an `alpha`/shape guard);
- an optional additive smoothing strength `lambda_smooth` (default `0`, i.e.
  exact maximum likelihood) shrinks each probability update towards its initial
  value, e.g. $m_s = (\sum q\,\mathbf{1} + \lambda\, m_s^{0}) / (\sum q +
  \lambda)$, which helps with sparse categorical levels;
- `NA` comparison values are dropped per field during fitting and contribute a
  neutral factor (log-density `0`) in scoring.

**Multistart (`nstart`).** With `nstart = 1` (the default) a single deterministic
data-driven start is used and the result is fully reproducible. With
`nstart > 1` the estimator runs that start first, then `nstart - 1` *perturbed*
starts (log-normal jitter of the parameters), and keeps the fit with the largest
observed log-likelihood. The perturbations use the package RNG and **respect any
`set.seed()`** (the RNG state is saved and restored on exit, leaving the caller's
stream untouched), so multistart fits are reproducible. `nstart >= 10` is
recommended in the pure-continuous regime, where the gamma mixture can have local
optima.

**Label-switching guard.** A mixture is invariant to relabelling its components,
so after convergence the estimators relabel so that **component 1 is always the
match**. The decision is *discriminative-power-weighted*: each field votes by the
size of its match/non-match separation (binary $|m - u|$; categorical, mass on
the exact-agreement level; continuous, the hurdle mass $p_0$ and the mean
distance $\alpha\beta$). Weighting by separation means an uninformative field
(where $m \approx u$) cannot flip the labels; on an exact tie the smaller mixing
component is taken as the match.

**Auto-detection and overrides.** `problink_em_mixed()` inspects each comparison
column and assigns a type: values in $\{0, 1\}$ become `binary`; non-negative
integers with few distinct values (at most `maxlevels`, default `10`) become
`categorical`; everything else becomes `continuous`. The inferred types are
reported via `message()`. The one genuinely ambiguous case is an integer-valued
*distance* with few distinct levels (e.g. a binned date lag), which detects as
`categorical`; pass `types =` to force it to `continuous` (or any field to any
type).

**Drop-in scoring.** Both `predict` methods compute, for each field, the
per-field log density ratio $\log(d_M / d_U)$ by block (Bernoulli, multinomial,
or hurdle-gamma), sum them into the matching weight $w_{ij}$, and form the
posterior $q_{ij} = m\,p_M / (m\,p_M + u\,(1 - p_M))$ from the density products.
Crucially they return the **same columns** as `predict.problink_em`
(`weights`, or `mprob`/`uprob`/`mpost`/`upost`). Consequently
`select_threshold`, `select_greedy`, `select_n_to_m`, `select_unique` and `link`
all work without any changes -- only the *estimator* differs from the classic
workflow.

## Notes and diagnostics

- **`p0` auto-derivation (blocking-aware).** When you do not supply `p0`, the
  EM start for the match prevalence is taken from the pairs object as
  `min(nA, nB) / nrow(pairs)` -- the largest plausible match rate (every record
  in the smaller file matches). Because `nrow(pairs)` already reflects any
  blocking, the start adapts to how many pairs you generated. It is only a
  *starting* value: the prevalence is re-estimated by the algorithm. Supply an
  explicit `p0` to override.

- **No exact matches is fine.** In the pure-continuous regime every `p0`
  converges to its floor `1/N`; `print()` reports the hurdle as *inactive* and
  the model is a plain gamma mixture. This is the correct MLE, and matches are
  recovered from the gamma-component separation. (Use `nstart >= 10` here, as
  shown above.)

- **Diagnostics.** Both `print()` methods report the fitted parameters, the
  initial and fitted match prevalence, the number of EM iterations and whether
  the algorithm converged; `summary()` adds the mean distances and
  log-likelihood. `problink_em_gamma` additionally reports the hurdle activity
  flag, and `problink_em_mixed` lists the inferred per-field types.

- **Drop-in output.** `predict()` for both models returns the same columns as
  `predict.problink_em` (`weights`, or `mprob`/`uprob`/`mpost`/`upost` via
  `type = "probs"`/`"all"`). Consequently `select_threshold`, `select_greedy`,
  `select_n_to_m`, `select_unique` and `link` all work without any changes --
  only the *estimator* differs from the classic workflow.

## References

Fellegi, I. P. and A. B. Sunter (1969). "A Theory for Record Linkage."
*Journal of the American Statistical Association* **64**(328): 1183-1210.

Vo, T. K. D., G. Chauvet, A. Happe, E. Oger, S. Paquelet and V. Garès (2023).
"Extending the Fellegi-Sunter record linkage model for mixed-type data."
*Computational Statistics & Data Analysis* **179**: 107656.
